---
title: "Tutorial for the samplyzer R library"
output:
  html_notebook: default
  html_document: default
---

This is a [R Markdown](http://rmarkdown.rstudio.com) Notebook for utilzing functions of the genotypeQC R library. When you execute code within the notebook, the results appear beneath the code.

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*.

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

### Install required libraries
This library required R version >= 3.3.2.
```{r}
#install.packages('devtools')
library(devtools)
install_github('xiaolicbs/samplyzer')
library(samplyzer)
```

### Load input files
```{r}
setwd('/Users/xiaoli/samplyzer/vignettes/data')
bamQcMetr = read.csv('bamQcMetr.tsv', sep = '\t')
vcfQcMetr = read.csv('vcfQcMetr.tsv', sep = '\t')
annotations = read.csv('sampleAnnotations.tsv', sep = '\t')
samplepc = read.csv('samplePCs.tsv', sep = '\t')
refpc = read.csv('refPCs.tsv', sep ='\t')
stratify = c('ANCESTRY', 'SeqTech')
```

### Construct an Sample Dataset object
```{r}
sds = sampleDataset(bamQcMetr = bamQcMetr, vcfQcMetr = vcfQcMetr, 
                    annotations = annotations, primaryID = 'SampleID')
```

### Attributes of sample data set
A Sample Data Set object contains several attributes, including a data frame, which contains all the data for samples and
several other attributes that contains metadata, such as qcMetrics, bamQcMetr, vcfQcMetr and PrimaryID. To check attributes of your current SDS:
```{r}
attributes(sds)
```
### Modify attributes of sample data set by
Function `setAttr` can be used to modify attributes of SampleDataset. For example, to add genotype PCs to the SampleDataset:
```{r}
sds = setAttr(sds, attributes = 'PC', data = samplepc, primaryID = 'SampleID')
```

### Access attributes of a SampleDataset
Function 'getAttr' was designed to access attributes of a SampleDataset, for example, you can access sample annotations via:
```{r}
getAttr(sds, 'PC', showID = T)
```
To know how many attributes are there, use:
```{r}
attributes(sds)
print(sds)
```
### Infer ancestry from
```{r}
sds = inferAncestry(sds, trainSet = refpc[,c('PC1', 'PC2', 'PC3')], knownAncestry = refpc$group )
getAttr(sds, 'inferredAncestry', showID = T)
```

### Perform a stratified QC metrics analysis
```{r}
sds = calZscore(sds, strat = c('inferredAncestry', 'SeqTech'), qcMetrics = sds$vcfQcMetr)
```

### Filter samples with hard cutoffs and z-scores
```{r}
cutoffs = data.frame(
  qcMetrics = c('Percent_lt_30bp_frags', 'Contamination_Estimation', 'Percent_Chimeric_Reads'),
  value = c(0.01, 0.02, 0.05),
  greater = c(T, T, T),
  stringsAsFactors = F
)
sds = flagSamples(sds, cutoffs = cutoffs, zscore = 4)
getAttr(sds, 'flaggedReason', showID = T) # show samples flagged
```

### Save SampleDataset
A S3 class function `save` is used to save SampleDataset to different formats.
```{r}
prefix = 'examples'
save(sds, RDS = paste(prefix, 'RDS', sep = '.'),
     tsv = paste(prefix, 'tsv', sep = '.'),
     xls = paste(prefix, 'xls', sep = '.'))
```

### Visualize a QC metric by annotations
```{r}
plt = sampleQcPlot(sds, annotation = 'inferredAncestry', qcMetrics = 'nHets', geom = 'scatter',
             main = 'nHet by inferredAncestry', outliers = 'Sample-001', show = T)

nhet = sampleQcPlot(sds, annotation = 'inferredAncestry', qcMetrics = 'nHets', geom = 'scatter',
             main = 'nHet by inferredAncestry', outliers = 'Sample-040', show = T)
```
```{r}
plots = PCplots(sds)
```
### Generate multipanel plots for
```{r}
grobList = sampleQcPlot(sds, qcMetrics = sds$bamQcMetr, annotation = 'LCSETMax', geom = 'scatter', ncols = 3, outliers = 'Sample-001')
ggplot2::ggsave('QcPlot.pdf', grobList, width = 12, height = 6)
```


### Interactive exploration of QC metrics
A shiny app was designed to interactively explore QC metrics
```{r}

```
