---
title: "Tutorial for the genotypeQC R library"
output: html_notebook
---

This is a [R Markdown](http://rmarkdown.rstudio.com) Notebook for utilzing functions of the genotypeQC R library. When you execute code within the notebook, the results appear beneath the code. 

Try executing this chunk by clicking the *Run* button within the chunk or by placing your cursor inside it and pressing *Cmd+Shift+Enter*. 

Add a new chunk by clicking the *Insert Chunk* button on the toolbar or by pressing *Cmd+Option+I*.

When you save the notebook, an HTML file containing the code and output will be saved alongside it (click the *Preview* button or press *Cmd+Shift+K* to preview the HTML file).

### Install required libraries
```{r}
install.packages('devtools')
install_github('xiaolicbs/genotypeqc')

library(devtools)
library(genotypeqc)
```

### Load input files
```{r}
setwd('~/genotypeqc/examples')
bamQcMetr = read.csv('bamQcMetr.tsv', sep = '\t')
vcfQcMetr = read.csv('vcfQcMetr.tsv', sep = '\t')
attributes = read.csv('sampleAttributes.tsv', sep = '\t')
stratify = c('ANCESTRY', 'SeqTech')
```

### Construct an Sample Dataset object
```{r}
sds = sampleDataset()
```
### Perform a stratified QC metrics analysis
```{r}
sds = calZscore(sds, strat = 'seqPcrEth', qcMetrics = sds$qcMetrics)
```

### Filter samples with hard cutoffs and z-scores
```{r}
sds = flagSamples(sds, cutoffs = cutoffs, zscore = 4)
```
### Add genotype PCs to the sample dataset object 
```{r}

```
